---
title: "Visualize Data"
author: "River Mckegney"
format: gfm
---


```{r}
#| label: setup
#| include: false
library()
```

## Create active Repository On Github: Here in Rstudio

> usethis::use_git()

'There are # uncommitted files:'
'! Is it ok to commit them?'
Selection: # (Definitly)
'Restart now?'
Selection: # (Yes)

> usethis::use_github()


## Your Turn 0

Add a setup chunk that loads the tidyverse packages.

```{r}
#| label: Load packages

library(tidyverse)
```


```{r}
#| label: Car Dataset

# head() function creates a small display containing first and last data sections
head(mpg)
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```


## Your Turn 2

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = hwy))
```


## Your Turn 3

Make a histogram of the `hwy` variable from `mpg`. Hint: do not supply a y variable.

```{r}
#| label: histogram

ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = hwy))
```

## Your Turn 4

Use the help page for `geom_histogram` to make the bins 2 units wide.

```{r}
#| label: histogram binwidth adjust

ggplot(data = mpg) +
  geom_histogram(mapping = aes(hwy), 
                 binwidth = 2)
```


## Your Turn 5

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy),
             size = 2,
             shape = 21,
             alpha = 0.5,
             color = "orange",
             fill = "blue"
             ) +
  labs(x = "Engine Displacement (L)",
       y = "Highway Miles Per Gallon", 
       color = "?")

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, 
                           colour = class)) +
  labs(x = "Engine Displacement (L)",
       y = "Highway Miles Per Gallon", 
       color = "Vehicle Type",
       title = "Effect of Engine Size on MPG") +
  theme_light()
```

## Help Me

What do `facet_grid()` and `facet_wrap()` do? 
(run the code, interpret, convince your group)

```{r}
# Makes a plot that the commands below will modify
q <- ggplot(mpg) + geom_point(aes(x = displ, y = hwy))

# Generates an engine displacement vs highway mpg graph for each category of vehicle cylinder type. Graphs are stacked and value increments matched to a standard grid so comparisons can easily be made.
q + facet_grid(. ~ cyl) 

# Creates graph sub-types for each category of vehicle drive type (front- rear-, and 4-wheel drive)
q + facet_grid(drv ~ .)

# Splits graphs by vehicle drive type and cylinder size
q + facet_grid(drv ~ cyl)

# Makes unique graphs split by each car class. Axis value labels are placed or "wrapped" around the axis margins of outermost graphs
q + facet_wrap(~ class)

# Also splits graph by car class, but all graphs are symmetrically layed across the grid
q + facet_grid(~ class)
```


## Your Turn 6

Make a bar chart `class` colored by `class`. Use the help page for `geom_bar` to choose a "color" aesthetic for class.

```{r}

```

## Quiz

What will this code do?

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy)) 
  ggsave("example.jpg")
```


***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```